---
title: 'STOR 455 Group Project (Due 5pm on April 19th)'
subtitle: "Bowser's sole Koopa Troopa"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
  word_document: default
---

Cameron Kaplinger (730570002)

# The Prediction (Required)
Our prediction of the cumulative box office of "The Super Mario Bros. Movie" by May 5 is $821,996,057  

# Summary of Justification (Required)

I got the data set from boxofficemojo where I was able to see the domestic box office data of The Super Mario Bros. Movie from 4/5/2023-4/16/2023. After that, I was able to create my simple linear model and find the predicted cumulative box office by multiplying the slope by 30 and by adding the intercept. 

# Data (Optional)
```{r}
library(readr)
library(polynom)
library(ggplot2)

Data = read_csv("Mario - Sheet1.csv")
head(Data)
```


```{r}
SimpleModel = lm(`Domestic Gross`~Days, data = Data)

Plot = ggplot(Data, aes(Days, `Domestic Gross`)) + geom_point()
Plot + stat_smooth(method = "lm", formula = y ~ poly(x,1), se = FALSE)
```


# Analysis (Optional)
```{r}
cor(Data$`Domestic Gross`,Data$Days)

PredictedGross = SimpleModel$coefficients[1] + (SimpleModel$coefficients[2] * 30)
PredictedGross
```


# References (Optional)

The data: https://www.boxofficemojo.com/release/rl1930593025/?ref_=bo_di_table_1