---
title: 'STOR 455 Homework #1'
subtitle: "20 points - Due Thursday 1/26 at 12:00pm"
output:
  pdf_document: default
  word_document: default
---

__Directions:__  This first assignment is meant to be a brief introduction to working with R in RStudio. You may (and should) collaborate with other students. If you do so, you must identify them on the work that you turn in. You should complete the assignment in an R Notebook, including all calculations, plots, and explanations. Make use of the white space outside of the R chunks for your explanations rather than using comments inside of the chunks. For your submission, you should knit the notebook to PDF and submit the file to Gradescope.

__Eastern Box Turtles:__ The Box Turtle Connection is a long-term study anticipating at least 100 years of data collection on box turtles. Their purpose is to learn more about the status and trends in box turtle populations, identify threats, and develop strategies for long-term conservation of the species. Eastern Box Turtle populations are in decline in North Carolina and while they are recognized as a threatened species by the International Union for Conservation of Nature, the turtles have no protection in North Carolina. There are currently more than 30 active research study sites across the state of North Carolina. Turtles are weighed, measured, photographed, and permanently marked. These data, along with voucher photos (photos that document sightings), are then entered into centralized database managed by the NC Wildlife Resources Commission. The _Turtles_ dataset (found under "Resources" on Sakai) contains data collected at The Piedmont Wildlife Center in Durham. 
```{r}
library(readr)

Turtles <- read_csv("Turtles.csv")

head(Turtles)
```

1)    The _Annuli_ rings on a turtle represent growth on the scutes of the carapace and plastron. In the past, it was thought that annuli corresponded to age, but recent findings suggest that this is not the case. However, the annuli are still counted since it may yield important life history information. Construct a least squares regression line that predicts turtles' _Annuli_ by their _Mass_.
```{r}
lsrl = lm(Annuli~Mass, data = Turtles)
lsrl
```


2)    Produce a scatterplot of this relationship (and include the least squares line on the plot).
```{r}
plot(Annuli~Mass, data = Turtles)
abline(lsrl)
```

3)    The turtle in the 40th row of the _Turtles_ dataset has a mass of 390 grams. What does your model predict for this turtle's number of _Annuli_? What is the residual for this case?
```{r}
Predicted = Turtles$Mass * lsrl$coefficients[2] + lsrl$coefficients[1]
Residual = Turtles$Annuli - Predicted

Predicted[40]
Residual[40]
```


4)    Which turtle (by row number in the dataset) has the largest positive residual? What is the value of that residual?
```{r}
largest_negative = min(Residual)
largest_negative
```


5)    Which turtle (by row number in the dataset) has the most negative residual? What is the value of that residual?
```{r}
largest_positive = max(Residual)
largest_positive
```


6)    Comment on how each of the conditions for a simple linear model are (or are not) met in this model. Include at least two plots (in addition to the plot in question 2) - with commentary on what each plot tells you specifically about the appropriateness of conditions.
```{r}
plot(lsrl, 1:2)
# The red line is around 0 and is relatively flat. The conditions for the linear model are met.
# The QQ plot shows the same thing. There are a lot of points on the line or close to the line.
```


7)    Experiment with at least two transformations to determine if models constructed with these transformations appear to do a better job of satisfying each of the simple linear model conditions. Include the summary outputs for fitting these models and scatterplots of the transformed variable(s) with the least square lines.
```{r}
log_model = lm(log(Annuli)~log(Mass), data = Turtles)
summary(log_model)
```


```{r}
exp(-0.37469)
```


```{r}
plot(log(Annuli)~log(Mass), data = Turtles)
abline(log_model)

plot(log_model, 1)

plot(Annuli~Mass, data = Turtles)
curve(0.6875024*(x^0.55594), add=TRUE)
```


8)    For your model with the best transformation from question 7 (It still may not be an ideal model), plot the raw data (not transformed) with the model (likely a curve) on the same axes.
```{r}
plot(Annuli~Mass, data = Turtles)
curve(0.6875024*(x^0.55594), add=TRUE)
```


9)    Again, the turtle in the 40th row of the _Turtles_ dataset has a mass of 390 grams. For your model using the best transformation from question 7, what does this model predict for this turtle's number of _Annuli_? In terms of _Annuli_, how different is this prediction from the observed value?
```{r}
log_annuli = 0.6875024*(390^0.55594)
difference = Turtles$Annuli[40] - log_annuli

log_annuli
difference
```


10)   For your model using the best transformation from question 7, could the relationship between _Mass_ and _Annuli_ be different depending on the _LifeStage_ and _Sex_ of the turtle? Construct two new dataframes, one with only adult male turtles, and one with only adult female turtles. Using your best transformation from question 7, construct two new models to predict _Annuli_ with _Mass_ for adult male and adult female turtles separately. Plot the raw data for _Anulli_ and _Mass_ for all adult turtles as well as each of these new models on the same plot. You should use different colors for each model (which are likely curves). What does this plot tell you about the relationship between _Mass_ and _Annuli_ depending on the _Sex_ of adult turtles?
```{r}
Male_turtles = Turtles[Turtles$Sex == "Male",]
Female_turtles = Turtles[Turtles$Sex == "Female",]

Male_turtles
Female_turtles
```

```{r}
male_log_model = lm(log(Annuli)~log(Mass), data = Male_turtles)
summary(male_log_model)

female_log_model = lm(log(Annuli)~log(Mass), data = Female_turtles)
summary(female_log_model)
```


```{r}
plot(Annuli~Mass, data = Male_turtles)
curve(exp(1.26995)*(x^0.27945), add=TRUE, col=c("blue", "blue"))

plot(Annuli~Mass, data = Female_turtles)
curve(exp(-0.20412)*(x^0.52467), add=TRUE, col=c("pink", "pink"))
# Annuli is more correlated with mass for females compared to males. 
# This is proven on the graphs where the females have less variance compared to the males.
```